---
title: "RmarkdownNguyenFinalProject"
author: "Grant Nguyen"
date: "11/1/2020"
output: github_document
---

```{r include = FALSE}
knitr::opts_chunk$set(message = FALSE)
```

# A Deeper Dive into GHG Emissions by Grant Nguyen

## Introduction

This project will focus on the data by CAIT countries on GHG emissions by countries, gases, and sectors of the economy. Most countries are include except China, which refused to partake. From my end, I wanted to flesh out and explore this data by visualizing many graphs, while also answering a few prominent questions. These questions are: Is the average greenhouse gas emissions from CAIT (China excuded) countries going down? Do individual greenhouse gasses follow the trend of GHG's as a whole and what is the percentage of each individual gas emission relative to all gas emissions? What are the GHG emissions relative to economic sectors and what is the breakdown for USA's emissions? Answering these question will help my report elucidate which gases are prominent and which sectors are the heavy emitters. 

Packages
```{r}
library(tidyverse)
library(zoo)
install.packages("RcppRoll")
library(RcppRoll)
library(lubridate)
```
Loading in data 
```{r}
countryGHG <- read.csv("https://raw.githubusercontent.com/espm-157/Nguyen-Grant-Final-Project/master/cait_annex_ghgemissions.csv")
countryGHG
```

## Renaming data
```{r}
countryGHG0 <- rename(countryGHG, "totalGHGemissions" = "Total.GHG.Emissions.Excluding.LUCF..MtCO2e.", "totalGHGemissionsLUCF" = "Total.GHG.Emissions.Including.LUCF..MtCO2e.", "totalCO2" = "Total.CO2..excluding.LUCF...MtCO2e.", "totalCH4" = "Total.CH4..MtCO2e.", "totalN2O" = "Total.N2O..MtCO2e.", "totalFgas" = "Total.F.Gas..MtCO2e.", "energy" = "Energy..MtCO2e.", "industrialprocesses" = "Industrial.Processes..MtCO2e.", "Solvent" = "Solvent.and.Other.Product.Use..MtCO2e.", "agriculture" = "Agriculture..MtCO2e.", "waste" = "Waste..MtCO2e.", "LUCF" = "Land.Use.and.Forestry..Net.Forest.Conversion...MtCO2e.", "energyindustries" = "Energy.Industries..MtCO2e.", "ManuIndusConst" = "Manufacturing.Industries.and.Construction..MtCO2e.", "transport" = "Transport..MtCO2e.", "othersectors" = "Other.Sectors..MtCO2e.", "energyother" = "Energy...Other..MtCO2e.", "fugitiveemissions" = "Fugitive.Emissions.from.Fuels..MtCO2e.")
```

## Exploring GHG Emissions
```{r}
countryGHG1 <- countryGHG0 %>%
  group_by(Year, Country) 
countryGHG1

G1 <- countryGHG1 %>% 
  ggplot(aes(x = Year, y = totalGHGemissions, color = Country)) +
  geom_line() + 
  ggtitle("GHG Emissions by Year") + 
  labs(y="Total GHG Emissions", x = "Year")
G1

ghg <- countryGHG1 %>% group_by(Year) %>% mutate(meanghg = mean(totalGHGemissions), na.rm=TRUE)

G2 <- ghg %>% 
  ggplot(aes(x = Year, y = meanghg, color = Country)) +
  geom_line() + 
  ggtitle("Average GHG Emissions by Year") + 
  labs(y="Avg GHG Emissions", x = "Year")
G2
```
### Analysis
I wanted to graph total GHG emissions by country to get a visual representation on which countries are emitting the most. There is a large amount of countries emitting lower than the top six emitters: USA, EU(15), EU(27), Russian Federation, Japan, and Germany. Note that EU(15) and EU(27) make up 15 and 17 countries. Taking that into account the top four emitters are USA, Russian Federation, Japan, and Germany. Additionally, I explored average GHG emissions by year to see if GHG emissions rate is decreasing. Form thr graph, it looks like the average GHG emission rate is moderately decreasing. 

## Breakdown of most commonly emitted GHGs
```{r}
percent <- countryGHG1 %>% mutate(percentco2 = totalCO2 / totalGHGemissions)

percent %>% 
  ggplot(aes(x = Year, y = percentco2, color = Country)) +
  geom_line() + 
  ggtitle("Percent CO2 Emissions by Year")

percent3 <- percent %>% group_by(Country) %>% summarize(mean = mean(percentco2), na.rm=TRUE)
percent3

percent3 %>% summarize(mean1 = mean(mean), na.rm=TRUE)

```

```{r}
percent1 <- countryGHG1 %>% mutate(percentch4 = totalCH4 / totalGHGemissions)

percent1 %>% 
  ggplot(aes(x = Year, y = percentch4, color = Country)) +
  geom_line() + 
  ggtitle("Percent CH4 Emissions by Year")

percent2 <- percent1 %>% group_by(Country) %>% summarize(mean = mean(percentch4), na.rm=TRUE)
percent2

percent2 %>% summarize(mean1 = mean(mean), na.rm=TRUE)
  
```

```{r}
percent4 <- countryGHG1 %>% mutate(percentn2o = totalN2O / totalGHGemissions)

percent4 %>% 
  ggplot(aes(x = Year, y = percentn2o, color = Country)) +
  geom_line() + 
  ggtitle("Percent N2O Emissions by Year")

percent5 <- percent4 %>% group_by(Country) %>% summarize(mean = mean(percentn2o), na.rm=TRUE)
percent5

percent5 %>% summarize(mean1 = mean(mean), na.rm=TRUE)
  
```

```{r}
percent6 <- countryGHG1 %>% mutate(percentfgas = totalFgas / totalGHGemissions)

percent6 %>% 
  ggplot(aes(x = Year, y = percentfgas, color = Country)) +
  geom_line() + 
  ggtitle("Percent F Gas Emissions by Year")

percent7 <- percent6 %>% group_by(Country) %>% summarize(mean = mean(percentfgas), na.rm=TRUE) 

percent7 <- na.omit(percent7)

percent7 %>% summarize(mean1 = mean(mean), na.rm=TRUE)
  
```
### Analysis
As you can see from the graphs that display an individual gas emission by year and country, the top GHG emitters highlighted before doesn't always correlate to the high gas percentage relative to their total emissions. To dive deeper into a pattern I was seeing. I calculated the average percentage of each gas's emission relative to other gas emissions by country and absolute average. The breakdown is 79.22% co2, 11.26% ch4, 8.22% n2o, and 1.44% f gas. Most of the GHG emission is of co2, followed not closely by ch4 and n2o.

##GHG by Sector of the Economy
```{r}
G6 <- countryGHG1 %>% 
  ggplot(aes(x = Year, y = energy, color = Country)) +
  geom_line() 
G6
```

```{r}
G7 <- countryGHG1 %>% 
  ggplot(aes(x = Year, y = industrialprocesses, color = Country)) +
  geom_line() 
G7
```

```{r}
G8 <- countryGHG1 %>% 
  ggplot(aes(x = Year, y = agriculture, color = Country)) +
  geom_line() 
G8
```

```{r}
G9 <- countryGHG1 %>% 
  ggplot(aes(x = Year, y = waste, color = Country)) +
  geom_line() 
G9
```
###Analysis
Here I visualize total GHG emissions by sectors of the economy. The main sectors of the economy I explored include: energy, industrial processes, agriculture, and waste. In general the four highest GHG emitter pointed out in the first section are the high emitters relative to each sector. It is interesting to note that the waste sector acts as a carbon sink according to data. It aborbs greenhouse gases. In addition the high emitters have also the highest carbon sink within the waste sector.

## US GHG Emissions and Sector Breakdown
```{r}
USA <- countryGHG0 %>% filter(Country == "United States of America")
USA

USA1 <- USA %>% 
  ggplot(aes(x = Year, y = totalGHGemissions)) +
  geom_line() + 
  ggtitle("USA's GHG Emission by Year") + 
  labs(y="Total GHG Emissions", x = "Year")
USA1

USA2 <- USA %>% mutate(meanenergy = mean(energy), meanip = mean(industrialprocesses), meansolvent = mean(Solvent), meanagri = mean(agriculture), meanwaste = mean(waste), meanlucf = mean(LUCF), na.rm=TRUE) %>% mutate(ghgsector = meanenergy + meanip + meansolvent + meanagri + meanwaste + meanlucf)
USA2

USA2 %>% summarise(meanenergy / ghgsector, meanip / ghgsector, meansolvent / ghgsector, meansolvent / ghgsector, meanagri / ghgsector, meanwaste / ghgsector, meanlucf / ghgsector)
``` 
###Analysis
From the total GHG emissions graph by the US, it clear that GHG emissions is increasing by the US. This is a worrying trend that counters the growing climate awareness trend in the 21st century. Additionally I wanted to give a breakdown of sector emissions within the US. I calculated that the energy sector has 98% of the GHG emissions, while agriculture has 7.57%, followed by industrial processes at 5.47%, land use at 2.46%, and solvent at 0.076%. Note that waste was negative and wasn't calculated in the percentage of emissions by sector. Technically, it's the most green sector.

## Wrap up
In general, this project help elucidate the data on GHG emissions by country and sector. Moving forward this data could be used by policy makers to hihglight high emitting sectors like energy and agriculture. It can also be used to point out how co2, ch4 and n2o are the most commonly emitted GHG, from where policy makers can help push for fundings torwards technical innovation that will help provide methods to reduce these gasses in the atmosphere. I had a enjoyable time applying what I learned from class projects to my own individual project.






